\myTitle

### 2024-02-12 body-text-of-paper.tex
Eric edited through line 262 in body-text-of-paper.tex

A best practice for evaluating the behavior of genetic clustering algorithms on an empirical dataset is to conduct
parallel analyses on data that have been simulated---so the truth is known---so as to mirror the empirical data set
as closely as possible \citep{vaha2006efficiency,anderson2008improved,latch2011fine}.
Techniques to simulate individual genotypes in this context frequently use sampling with replacement from the
observed allele frequencies (see, for example, \citealt{nielsen2006hybridlab,kinziger2008hybridization}).
However, sampling with replacement implicitly increases the pairwise genetic distance among simulated samples, relative to the observed
samples, and
inflates the perceived statistical power to correctly assign individuals back to source populations.
We refer to this phenomenon as resampling-induced, spurious power inflation (RISPI).
\citet{anderson2008improved} used simulations to document this type of artifact in the context of
genetic stock identification (GSI), a type of clustering application in which individuals are constrained to have
all their ancestry from a single subpopulation.
Similar statistical artifacts occur when simulating data to
assess how accurately admixture fractions can be estimated.

Beyond simply evaluating the resolution of discrete populations in a simulated framework, researchers may also be
interested in how individuals of admixed ancestry are genetically characterized given the statistical power of the available dataset and the
analysis methods used.
Recognizing that connectivity among genetically distinct populations may be relatively rare, additional insight into the
frequency of immigration can be gained by integrating across generations to identify both direct migrants and descendants of migrants
 (i.e., F1 hybrids or advanced hybrids reflecting the effect of backcrossing across multiple generations [BC1, \ldots,
BCX]).
Evaluating the results of genetic clustering algorithms in the context of data that are
expected to arrive by such a process of hybridization requires simulation of
explicit hybrids.  Such simulations are the purpose of the software {\sc hybridlab} 
\citep{nielsen2006hybridlab}.

@@

{\sc hybridlab} was conceived at a time when
the data sets used in molecular ecology universally included very few markers, such that
it was reasonable to assume the markers were unlinked, and, hence, independently segregating.
Today, however, with data sets of thousands of markers, that assumption is badly violated---alleles
on the same chromosome are likely to be segregated together into the next generation.
Simulating linked loci as independent creates simulated hybrids, beyond the $F1$ category, with 
artificially low variance in admixture fraction, leading to the false interpretation that different
hybrid categories are more easily distinguished than they are. Therefore, in evaluating descendants of migrants within a simulated framework, it is imperative to take into consideration the pattern of
linkage among loci available for analysis. This was done recently to assess the power of a SNP panel for identifying descendants of escaped, farm-raised Atlantic salmon \citep{wringe2019development}; however the method used (described at https://eriqande.github.io/SalarHybPower/) was developed in an {\em ad hoc} fashion, was not easily extended to
complex hybridization scenarios, and was not at all user-friendly.

In response to these needs, we developed \gscramble{} – a simulation approach that samples alleles from respective populations without
replacement, thus maintaining genetic differences between the clusters commensurate to those in the empirical dataset, and eliminating the effect of RISPI.
Further, by simulating individual genotypes based on user-defined pedigrees of arbitrary complexity, \gscramble{} allows for the simulation of admixed genotypes resulting from any conceivable category of hybrid, while allowing the tracking of haplotype blocks from source populations within the specified pedigree.
By integrating species-specific recombination rates, \gscramble{} simulates biologically informed genotypes that mirror empirical data.
We develop and illustrate \gscramble{} with use of single nucleotide polymorphic (SNP) genotypes, though it is equally applicable to data sets with hundreds or thousands of
microhaplotypes \citep{baetscher2018microhaplotypes} or microsatellites
\citep{zhan2017megasat}.

@@

Before we proceed to a description of the methodology used by \gscramble{},
we demonstrate two important points through two simple simulations.  In the first,
we show that using a standard sampling-with-replacement
approach induces RISPI when estimating admixture fractions. In the second, we show
the effect of physical linkage on the distribution of admixture fractions of hybrid individuals,
which vividly shows why it is not acceptable with large numbers of markers, to simulate
the genotypes of hybrid individuals assuming no physical linkage.

@@

We also simulated the genomes of 48,000 individuals of each
hybrid category using \gscramble{}'s {\footnotesize\tt segregate()} function, as described
below, with number of chromosomes and sex-averaged recombination rates
for pigs, as estimated by from \citet{tortereau2012high}. In this case, we recorded the
admixture fraction of each simulated individual as the fraction of all genome segments
that originated from species~$a$.  This provides the distribution of the true admixture fraction
when physical linkage of markers is not ignored.

@@

Our scheme, described here, extends the traditional cross-validation approach
by simulating the descent of chromosomal linkage blocks through a user-specified
pedigree to create admixed individuals.  By imposing certain constraints
on the pedigree, our framework ensures that all the genetic material
within the original individuals is used to either create admixed individuals or is retained
within reference individuals (i.e., those that are purely of one
putative subopulation), and yet no new genetic material is
created by sampling with replacement from the original samples.

We refer to the structure used to implement this cross-validation simulation
of admixed individuals as
a {\em genomic simulation pedigree} or GSP.  Like any pedigree,
a GSP includes a set of founders (individuals that have no parents specified in the
pedigree), and it includes non-founders; however, we introduce an additional node
type in the GSP that represents samples taken from the non-founders. These samples
are the simulated admixed individuals.  A GSP must have no inbreeding loops, since
any inbreeding loop indicates that more than one copy of a gene in an ancestor may
be created amongst its descendants, which is a type of sampling with
replacement.  Furthermore, in a GSP, the amount of genetic material taken from the
samples at the sample nodes should be equal to the amount of material in the founders,
to ensure that all available data is being used to estimate power for the estimation of
admixture fractions.

>>>>

A best practice for evaluating the behavior of genetic clustering algorithms on an empirical dataset is to conduct
parallel analyses on data that have been simulated---so the truth is known---so as to mirror the empirical data set
as closely as possible \citep{vaha2006efficiency,anderson2008improved,latch2011fine}.
Techniques to simulate individual genotypes in this context frequently use sampling with replacement from the
observed allele frequencies (see, for example, \citealt{nielsen2006hybridlab,kinziger2008hybridization}).
However, sampling with replacement implicitly increases the pairwise genetic distance among simulated samples, relative to the observed
samples, and
inflates the perceived statistical power to correctly assign individuals (or segments of the genome) back to source populations.
We refer to this phenomenon as resampling-induced, spurious power inflation (RISPI).
\citet{anderson2008improved} used simulations to document this type of artifact in the context of
genetic stock identification (GSI), a type of clustering application in which individuals are constrained to have
all their ancestry from a single subpopulation.
Similar statistical artifacts occur when simulating data to
assess how accurately admixture fractions can be estimated.

Beyond simply evaluating the resolution of discrete populations in a simulated framework, researchers may also be
interested in how individuals of admixed ancestry are genetically characterized given the statistical power of the available dataset and the
analysis methods used.
Recognizing that connectivity among genetically distinct populations may be relatively rare, additional insight into the
frequency of immigration can be gained by integrating across generations to identify both direct migrants and descendants of migrants
 (i.e., F1 hybrids or advanced hybrids reflecting the effect of backcrossing across multiple generations $[BC_1, \ldots,
BC_X])$.
Evaluating the results of genetic clustering algorithms in the context of data that are
expected to arrive by such a process of hybridization requires simulation of
explicit hybrids.  Such simulations are the purpose of the software {\sc hybridlab} 
\citep{nielsen2006hybridlab}.

@@

{\sc hybridlab} was conceived at a time when
the data sets used in molecular ecology universally included very few markers, such that
it was reasonable to assume the markers were unlinked, and, hence, independently segregating.
Today, however, with data sets of thousands of markers, that assumption is badly violated---alleles
on the same chromosome are likely to be segregated together into the next generation.
Simulating linked loci as independent creates simulated hybrids, beyond the $F1$ category, with 
artificially low variance in admixture fraction, leading to the false interpretation that different
hybrid categories are more easily distinguished than they are. Therefore, in evaluating descendants of migrants within a simulated framework, it is imperative to take into consideration the pattern of
linkage among loci available for analysis. This was done recently to assess the power of a SNP panel for identifying descendants of escaped, farm-raised Atlantic salmon \citep{wringe2019development}; however the method used (described at https://eriqande.github.io/SalarHybPower/) was developed in an {\em ad hoc} fashion, was not easily extended to
complex hybridization scenarios, and is not user-friendly.

In response to these needs, we developed \gscramble{} – a simulation approach that samples alleles from respective populations without
replacement, thus maintaining genetic differences between the clusters commensurate to those in the empirical dataset, and eliminating the effect of RISPI.
Further, by simulating individual genotypes based on user-defined pedigrees of arbitrary complexity, \gscramble{} allows for the simulation of admixed genotypes resulting from any conceivable category of (non-inbred) hybrid, while allowing the tracking of haplotype blocks from source populations within the specified pedigree.
By integrating species-specific recombination rates, \gscramble{} simulates biologically informed genotypes that mirror empirical data.
We develop and illustrate \gscramble{} through the use of single nucleotide polymorphism (SNP) genotypes, though it is equally applicable to data sets with hundreds or thousands of
microhaplotypes \citep{baetscher2018microhaplotypes} or microsatellites
\citep{zhan2017megasat}.

@@

Before we proceed to a description of the methodology used by \gscramble{},
we demonstrate two important points through two simple simulations.  In the first,
we show that using a standard sampling-with-replacement
approach induces RISPI when estimating admixture fractions. In the second, we show
the effect of physical linkage on the distribution of admixture fractions of hybrid individuals,
which vividly shows why it is not acceptable, with large numbers of markers, to simulate
the genotypes of hybrid individuals assuming no physical linkage.

@@

We also simulated the genomes of 48,000 individuals of each
hybrid category using \gscramble{}'s {\footnotesize\tt segregate()} function, as described
below, with number of chromosomes and sex-averaged recombination rates
for pigs, as estimated by from \citet{tortereau2012high}. In this case, we recorded the
admixture fraction of each simulated individual as the cumulative length of all genome segments
that originated from species~$a$, divided by the total genome length.  This provides the distribution of the true admixture fraction
when physical linkage of markers is not ignored.

@@

Our scheme, described here, extends the traditional cross-validation approach
by simulating the descent of chromosomal segments through a user-specified
pedigree to create admixed individuals.  By imposing certain constraints
on the pedigree, our framework ensures that all the genetic material
within the original individuals is used to either create admixed individuals or is retained
within reference individuals (i.e., those that are purely of one
putative subopulation), and yet no new genetic material is
created by sampling with replacement from the original samples.

We refer to the structure used to implement this cross-validation simulation
of admixed individuals as
a {\em genomic simulation pedigree} or GSP.  Like any pedigree,
a GSP includes a set of founders (individuals that have no parents specified in the
pedigree), and it includes non-founders; however, we introduce an additional node
type in the GSP that represents samples taken from the non-founders. These samples
are the simulated admixed individuals.  A GSP must have no inbreeding loops, since
any inbreeding loop indicates that more than one copy of a gene in an ancestor may
be created amongst its descendants, which is a type of sampling with
replacement.  Furthermore, in a GSP, the amount of genetic material taken from the
samples at the sample nodes should be equal to the amount of material in the founders,
to ensure that all available data are being used in the simulations to predict the power for
estimating admixture fractions.
